QPLIB: A Library of Quadratic Programming Instances

### 2016-12-18 QPLIB-3.tex
Update QPLIB-3.tex
checked up to §3.2, will continue later

The first goal of the library is to represent the largest possible majority of
the different categories of QP problems.
The second goal of the library is to gather ``challenging'' instances,
i.e., instances which can not be easily solved by  state-of-the-art
solvers.
The third goal is to include in each of the categories a set of well
diversified instances.
Finally, the fourth goal is to obtain a library whose set of instances
is neither too small either too large, to preserve statistical
relevance while at the same time being computationally manageable.

To achieve the aforementioned goals, we performed the following two
filters applied in cascade.

\begin{itemize}

\item \emph{First Instances Filter.}\\
The first filter is design to drastically reduce the number of
instances by eliminating the ``easy'' ones. An empirical measure way
of testing the hardness of an instance is the CPU time needed by a
complete solver (cf.~\S \ref{sec:algo}) to solve the instance to
global optimality.
%n this step, we decided to use the broadest set of solvers able to tackle the different categories of quadratic problems.\\
Accordingly, for each of the gathered instance, we run the solvers in
{\tt Gams} (see Table \ref{t:solvers}) able to solve it to global optimality.
The number of solvers depends on the category of the instances under
consideration.
We then filtered according to a relative measure of computational
difficulty, i.e., we discarded all instances that are solved by at
least 30\% of the complete solvers within a time limit of 30 seconds.

\item \emph{Second Instances Filter.}\\
The goal of the second filter is to eliminate instances that are similar.
%We used the instance features presented in
%Section~\ref{subsec:feature} and the CPU solving time obtained in the
%first filter to guide this second step. \\
We carefully analysed the instances and we grouped them accordingly to
similar features. To reduce the size of the test-bed
while keeping the representativeness of the library,   we kept for
each only few instances.
Finally in order to have a deeper insight on the computational hardness of the
instances, we performed a second round of tests on the filtered set of
instances.  We used a general purpose solver able to tackle all QP categories and we set a time limit of 120 seconds. 
All the instances which have been solved to
proven optimality within this time limit have been discarded.
\end{itemize}
In Table~\ref{tab:filters} we summarize the two filter steps, which
allows us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

\begin{center}
\begin{table}[]
 \centering

 \setlength{\tabcolsep}{5pt}
%  \arraystretch{1}
\begin{tabular}{cccc}
Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 Instances }& \\
& \multicolumn{ 2}{c}{$\Downarrow$}& \\
& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. inst. & \\
First Filter  & $\Downarrow$  & $\Downarrow$ & \\
 & $\approx$ 3000 Discr. Inst.  & $\approx$ 1000 Cont. Inst. & \\
Second Filter & $\Downarrow$  & $\Downarrow$  & \\
% & 600 Discr. Inst.  & 250  Cont. inst. & \\
  & 251 Discr. Inst.  & 116  Cont. inst. & \\
\end{tabular}
%\begin{center}\end{center}
\caption{Instance filter steps} \label{tab:filters}
\end{table}
\end{center}

>>>>

During the development of the library, a discussion ensued about what
the expected gols that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S \ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in {\tt Gams} (cf.~Table \ref{t:solvers})
       capable of solving it, whose number depends on the category of the
       instance under consideration. We then filtered according to a relative
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives. Finally, in order to only keep computationally
       challenging instances we ran a general purpose solver able to tackle
       all QP categories with a time limit of 120 seconds; all the instances
       which have been solved to proven optimality within this time limit were
       discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

\begin{center}
\begin{table}[]
 \centering
 \setlength{\tabcolsep}{5pt}
%  \arraystretch{1}
\begin{tabular}{cccc}
Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 Instances }& \\
& \multicolumn{ 2}{c}{$\Downarrow$}& \\
& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. inst. & \\
First Filter  & $\Downarrow$  & $\Downarrow$ & \\
 & $\approx$ 3000 Discr. Inst.  & $\approx$ 1000 Cont. Inst. & \\
Second Filter & $\Downarrow$  & $\Downarrow$  & \\
% & 600 Discr. Inst.  & 250  Cont. inst. & \\
  & 251 Discr. Inst.  & 116  Cont. inst. & \\
\end{tabular}
%\begin{center}\end{center}
\caption{Instance filter steps} \label{tab:filters}
\end{table}
\end{center}
